### 2016-04-22 paper.tex
Initial

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}

\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{paralist}
\usepackage{mathtools}
\usepackage{dingbat}
\usepackage{subcaption}
\usepackage{balance}
\usepackage{gensymb}
\usepackage{marginnote}
\usepackage{adjustbox}

\makeatletter
\newcommand{\@BIBLABEL}{\@emptybiblabel}
\newcommand{\@emptybiblabel}[1]{}
\makeatother
\usepackage{hyperref}

\sloppy

% \aclfinalcopy % Uncomment this line for the final submission
% \def\aclpaperid{34} %  Enter the acl Paper ID here

\usepackage{color}
\newcommand{\todo}[1]{}
\renewcommand{\todo}[1]{{\color{red} TODO: {#1}}}

%\renewcommand{\baselinestretch}{0.95}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\title{On Similarity}

% \author{First Author \\
%   Affiliation / Address line 1 \\
%   Affiliation / Address line 2 \\
%   Affiliation / Address line 3 \\
%   {\tt email@domain} \\\And
%   Second Author \\
%   Affiliation / Address line 1 \\
%   Affiliation / Address line 2 \\
%   Affiliation / Address line 3 \\
%   {\tt email@domain} \\}

\date{}

\begin{document}

\maketitle

% \begin{abstract}
% % \input{abstract.tex}
% \end{abstract}

% \section{Introduction}
% \label{sec:introduction}

Word occurrence patterns define its meaning \cite{firth1957lingtheory}. Moreover, the difference in occurrence quantifies the difference in meaning \cite{harris1954distributional}. The field of distributional semantics studies how and to what extent word meaning can be captured \cite{Turney:2010:FMV:1861751.1861756}. The evaluation of the models is usually done by comparing model's estimates with human judgements of relatedness (e.g.,~MEN \newcite{Bruni:2012:DST:2390524.2390544}) or similarity (e.g., SimLex-999~\newcite{hill2014simlex}).

There is a number of flaws in the current evaluation methods, which this paper addresses.

Intuitively, similarity is the amount of common properties between the entities, and is implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}. However, the similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}). For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} relatively have much more in common than \textit{trick} and \textit{size}.

The presence of incompatible pairings such as \textit{trick} and \textit{size} can be motivated by an attempt to control for false positives in the case when the correlation between the model predictions and the human judgements is used as evaluation. For vector space models, a large amount of random pairings\footnotemark{} might influence a particular arrangements of the clusters of similar words, for example, the cluster of fruit names has to be between the cluster of ``furious adjectives'' and ``mass nouns'', even if such arrangement was not intended by the dataset creators.

\footnotetext{15\% of MEN entries have score less or equal to 10 out of 50 and by design 1000 word pairs are sampled from pairs that are assigned low cosine similarity scores by a text-based distributional model.}

Semantic models of meaning presuppose a special way of arranging words meaning focusing on the fact that words with similar meaning are close to each other, while it is less important how far and where are the words with dissimilar meanings. The evaluation sets should state this implicitly.

% remove publisher, month etc from conf proceedings:
% \bibliographystyle{acl-short}
\bibliographystyle{acl2016}
\bibliography{references,dmilajevs_publications}
\balance

\end{document}

### 2016-04-26 paper.tex
Day two.

@@
 
-%\renewcommand{\baselinestretch}{0.95}
+% \renewcommand{\baselinestretch}{0.95}
 
@@
 
-Word occurrence patterns define its meaning \cite{firth1957lingtheory}.
-Moreover, the difference in occurrence quantifies the difference in meaning \cite{harris1954distributional}.
-The field of distributional semantics studies how and to what extent word meaning can be captured \cite{Turney:2010:FMV:1861751.1861756}.
-The evaluation of the models is usually done by comparing model's estimates with human judgements of relatedness (e.g.,~MEN \newcite{Bruni:2012:DST:2390524.2390544}) or similarity (e.g., SimLex-999~\newcite{hill2014simlex}).
+Similarity plays a crucial role in psychological theories of knowledge and behavior, for example, it is used to classify objects \cite{Tversky1977}.
+%
+Linguistic treatment of semantic similarity is based on two ideas: word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
 
-There is a number of flaws in the current evaluation methods, which this paper addresses.
+Practically, this leads to systems that use similarity to perform natural language processing tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
 
-Intuitively, similarity is the amount of common properties between the entities, and is implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
-However, the similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
-For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} relatively have much more in common than \textit{trick} and \textit{size}.
+In recent years, similarity received a lot of attention as an intrinsic evaluation of meaning representations \cite{mikolov2013efficient,Bruni:2012:DST:2390524.2390544}, because in a complete text processing system it is difficult to measure performance of a single (similarity) component.
+As a result, a new generation of datasets appeared from the computational linguistics community, for example, MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}, that surpasses older datasets stemmed in psychology (refer to \newcite{1986-13502-00119860101} for a comprehensive list of datasets) and information retrieval (most notably \newcite{2002:PSC:503104.503110}) in quantity and attention to the evaluated relation.
 
-The presence of incompatible pairings such as \textit{trick} and \textit{size} can be motivated by an attempt to control for false positives in the case when the correlation between the model predictions and the human judgements is used as evaluation.
-For vector space models, a large amount of random pairings\footnotemark{} might influence a particular arrangements of the clusters of similar words, for example, the cluster of fruit names has to be between the cluster of ``furious adjectives'' and ``mass nouns'', even if such arrangement was not intended by the dataset creators.
+Despite improvement, current similarity datasets deviate from the psychological understanding of similarity.
+Intuitively, similarity is the amount of common properties between entities, and is implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
 
+Categories are determined in terms of focal points, for example fruit names are arranged around the word \textit{fruit} in such a way that it is their nearest neighbor \cite{1986-13502-00119860101}.
+It is extremely important for a good semantic model to capture this property.
+Unfortunately, word pair similarity judgements do not capture this, as it is possible to arrange all fruit names close to \textit{fruit} but \textit{fruit} would not be the most common nearest neighbor.
+Moreover, as \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from lexical ontologies such as WordNet \cite{Miller:1995:WLD:219717.219748}, for example, \textit{traffic} and \textit{water} might be considered to be similar because of functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
+
+Moreover, the similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
+For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
+
+While the presence of incompatible pairings can be motivated as an attempt to control for false positives, they sound unnatural to human.
+Also, a large amount of random pairings\footnotemark{} might influence a particular arrangement of the clusters of similar words, for example, the cluster of fruit names has to be between the cluster of \textit{furious adjectives} and \textit{mass nouns}, even if such arrangement was not intended by the dataset creators.
 \footnotetext{15\% of MEN entries have score less or equal to 10 out of 50 and by design 1000 word pairs are sampled from pairs that are assigned low cosine similarity scores by a text-based distributional model.}
 
-Semantic models of meaning presuppose a special way of arranging words meaning focusing on the fact that words with similar meaning are close to each other, while it is less important how far and where are the words with dissimilar meanings.
-The evaluation sets should state this implicitly.
+Evaluation methods need to focus on how well a model is able to recover human similarity intuitions expressed as clusters (e.g~\textit{fruits}, \textit{company} and \textit{mass noun}).
+It can be seen as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements.
+
+This leads two points this paper addresses.
+The similarity judgements should be collected in a different way which reflects human intuition about similarity.
+Even though vector space models of meaning measure similarity between words, it should not be (and can not) be contrasted directly with human judgements, instead the model needs to recover the concept clusters that humans define.
 

### 2016-04-28 paper.tex
another revision

@@
 In recent years, similarity received a lot of attention as an intrinsic evaluation of meaning representations \cite{mikolov2013efficient,Bruni:2012:DST:2390524.2390544}, because in a complete text processing system it is difficult to measure performance of a single (similarity) component.
-As a result, a new generation of datasets appeared from the computational linguistics community, for example, MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}, that surpasses older datasets stemmed in psychology (refer to \newcite{1986-13502-00119860101} for a comprehensive list of datasets) and information retrieval (most notably \newcite{2002:PSC:503104.503110}) in quantity and attention to the evaluated relation.
+As a result, a new generation of similarity datasets appeared from the computational linguistics community.
+The two notable datasets are MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}.
+They surpasses older datasets stemmed in psychology (refer to \newcite{1986-13502-00119860101} for a comprehensive list of datasets) and information retrieval (for example \newcite{2002:PSC:503104.503110}) in quantity and attention to the evaluated relation.
+Both of them consist of word pairs assigned similarity (relatedness) scores based on human judgements.
 
-Despite improvement, current similarity datasets deviate from the psychological understanding of similarity.
+Despite improvement, these datasets deviate from the psychological understanding of similarity.
 Intuitively, similarity is the amount of common properties between entities, and is implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
 
-Categories are determined in terms of focal points, for example fruit names are arranged around the word \textit{fruit} in such a way that it is their nearest neighbor \cite{1986-13502-00119860101}.
+One of the uncaptured properties of human similarity judgements is that categories are formed around \emph{focal points} \cite{1986-13502-00119860101}.
+Consider fruit names assessed for similarity.
+They will be arranged around the word \textit{fruit} in such a way that it is their nearest neighbor, making \textit{fruit} the focal point of the category.
 It is extremely important for a good semantic model to capture this property.

### 2016-04-29 paper.tex

--- 2016-04-26 paper.tex
@@
 
-% \begin{abstract}
-% % \input{abstract.tex}
-% \end{abstract}
+\begin{abstract}
 
-% \section{Introduction}
-% \label{sec:introduction}
+\end{abstract}
 
-Similarity plays a crucial role in psychological theories of knowledge and behavior, for example, it is used to classify objects \cite{Tversky1977}.
-%
+\section{Introduction}
+\label{sec:introduction}
+
+Similarity plays a crucial role in psychological theories of knowledge and behaviour, for example, it is used to classify objects \cite{Tversky1977}.
+% TODO: maybe, mention \cite{Kriegeskorte2008}
+% TODO: a nicer transition
 Linguistic treatment of semantic similarity is based on two ideas: word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
@@
 Practically, this leads to systems that use similarity to perform natural language processing tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
+Because it is difficult to measure performance of a single (similarity) component in a pipliene, similarity datasets became popular among computational linguists to perform intrinsic evaluation of meaning representations.
 
-In recent years, similarity received a lot of attention as an intrinsic evaluation of meaning representations \cite{mikolov2013efficient,Bruni:2012:DST:2390524.2390544}, because in a complete text processing system it is difficult to measure performance of a single (similarity) component.
-As a result, a new generation of datasets appeared from the computational linguistics community, for example, MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}, that surpasses older datasets stemmed in psychology (refer to \newcite{1986-13502-00119860101} for a comprehensive list of datasets) and information retrieval (most notably \newcite{2002:PSC:503104.503110}) in quantity and attention to the evaluated relation.
+Two, currently widely used, datasets are MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}.
+They are designed especially for meaning representation evaluation and surpass datasets stemmed in psychology\footnote{Refer to \newcite{1986-13502-00119860101} for a comprehensive list of datasets.} and information retrieval \cite{2002:PSC:503104.503110} in quantity and attention to the evaluated relation.
 
-Despite improvement, current similarity datasets deviate from the psychological understanding of similarity.
-Intuitively, similarity is the amount of common properties between entities, and is implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
+Intuitively, similarity is the amount of common properties between entities, and can be implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
+The datasets control for this property by providing similarity (relatedness) scores between word pairs.
+%
+% TODO: the argument about focal points is still vague, because similarity scores still can express it.
+Maybe the argument should be given as a combination of centrality and ambiguity of low similarity.
+However, they overlook the following properties:
+\begin{compactitem}
+    \item Similarity scores group words to categories that are formed around \emph{focal points}, the terms that represent clusters.
+    % For this one needs to have all pairwise pairings or pairings between the focal point and the words of a category for each category.
+    \item Low similarity scores are ambiguous, they might mean dissimilarity between antonymy as between \textit{smart} and \textit{dumb} or absence of similarity in the case of \textit{trick} and \textit{size}.
+\end{compactitem}
 
-Categories are determined in terms of focal points, for example fruit names are arranged around the word \textit{fruit} in such a way that it is their nearest neighbor \cite{1986-13502-00119860101}.
+% TODO: wrap up the introduction and make a transition to the next sections.
+
+\section{Similarity properties}
+
+\paragraph{Centrality}
+
+It has been shown empirically that similarity judgements geometrically arrange objects in such a way that the categories are formed around \emph{focal points}.
+This means that similarity judgements given by humans arrange fruit names around the word \textit{fruit} in such a way that it is their nearest neighbour, making \textit{fruit} the focal point of the category of \textit{fruits} \cite{1986-13502-00119860101}.
+This property is referred to as \emph{centrality}.
+%
 It is extremely important for a good semantic model to capture this property.
-Unfortunately, word pair similarity judgements do not capture this, as it is possible to arrange all fruit names close to \textit{fruit} but \textit{fruit} would not be the most common nearest neighbor.
-Moreover, as \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from lexical ontologies such as WordNet \cite{Miller:1995:WLD:219717.219748}, for example, \textit{traffic} and \textit{water} might be considered to be similar because of functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
+Unfortunately, word pair similarity judgements do not capture this, as it is possible to arrange all fruit names close to \textit{fruit} but \textit{fruit} would not be the most common nearest neighbour.
 
-Moreover, the similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
+To complicate the matter, existing lexical databases such as WordNet \cite{Miller:1995:WLD:219717.219748} can not be used to control for centrality.
+As \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies defined in those databases.
+For example, \textit{traffic} and \textit{water} might be considered to be similar because of functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
+
+\paragraph{Ambiguity}
+
+Another issue is that the similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
 For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
@@
 While the presence of incompatible pairings can be motivated as an attempt to control for false positives, they sound unnatural to human.
-Also, a large amount of random pairings\footnotemark{} might influence a particular arrangement of the clusters of similar words, for example, the cluster of fruit names has to be between the cluster of \textit{furious adjectives} and \textit{mass nouns}, even if such arrangement was not intended by the dataset creators.
+Also, a large amount of random pairings\footnotemark{} might influence a particular arrangement of the clusters of similar words, for example, the cluster of \textit{fruits} has to be between the cluster of \textit{furious adjectives} and \textit{mass nouns}, even if such arrangement was not intended by the dataset creators.
 \footnotetext{15\% of MEN entries have score less or equal to 10 out of 50 and by design 1000 word pairs are sampled from pairs that are assigned low cosine similarity scores by a text-based distributional model.}
 
-Evaluation methods need to focus on how well a model is able to recover human similarity intuitions expressed as clusters (e.g~\textit{fruits}, \textit{company} and \textit{mass noun}).
-It can be seen as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements.
+\section{Dataset design}
 
-This leads two points this paper addresses.
+% Some of the properties:
+%
+% * items belong to groups, such as fruits, animals, etc.
+% * all items in the group are assessed for similarity
+% * similarity is on the score of -3...+3.
+Where negative items indicate dissimilarity (hot, cold: -3), positive indicate similarity (hot, warm: 2) and 0 is reserved for incompatible pairs (trick, size:0)
+
+% TODO: What if we ask not for similarity, but for confusion? Townsend (1971) "Theoretical analysis of an alphabetic confusion matrix".
+% TODO: Sorting seems to be a better fit! Rosenberg & Kim, (1975) "The Method of Sorting as a Data-Gathering Procedure in Multivariate Research".
+
+Evaluation methods need to focus on how well a model is able to recover human similarity intuitions expressed as groupings around their focal points (e.g~\textit{fruits}, \textit{companies} and \textit{mass nouns}).
+We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
+
 The similarity judgements should be collected in a different way which reflects human intuition about similarity.
@@
 
-% remove publisher, month etc from conf proceedings:
-% \bibliographystyle{acl-short}
+% TODO: discuss antonymy and incompatible notions.
+
 \bibliographystyle{acl2016}

### 2016-04-30 paper.tex

@@
 \usepackage{adjustbox}
-
-\makeatletter
-\newcommand{\@BIBLABEL}{\@emptybiblabel}
-\newcommand{\@emptybiblabel}[1]{}
-\makeatother
-\usepackage{hyperref}
 
@@
 
-Similarity plays a crucial role in psychological theories of knowledge and behaviour, for example, it is used to classify objects \cite{Tversky1977}.
-% TODO: maybe, mention \cite{Kriegeskorte2008}
-% TODO: a nicer transition
-Linguistic treatment of semantic similarity is based on two ideas: word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
+Similarity plays a crucial role in psychological theories of knowledge and behaviour.
+For example, it is used for classification and conceptualisation \cite{Tversky1977}.
+\textit{Fruit} is a \emph{category} because it is a useful generalisations in everyday life: fruits are sweet and constitute deserts, so when one is presented an unseen fruit, she can hypothesise that it is served towards the end of a dinner.
 
-Practically, this leads to systems that use similarity to perform natural language processing tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
-Because it is difficult to measure performance of a single (similarity) component in a pipliene, similarity datasets became popular among computational linguists to perform intrinsic evaluation of meaning representations.
+Generalisations are extremely powerful in describing a language as well.
+The verb \textit{runs} requires its subject to be singular.
+\textit{Verb}, \textit{subject} and \textit{singular} are categories that are used to describe English grammar.
+When one encounters an unknown word and is told that it is a verb, he will immediately have an idea how to use it assuming that it is used similarly to other English verbs.
+
+Linguistic formalisation of similarity could base on two ideas: word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
+From computational perspective, this motivates and guides development of similarity components that are embedded into natural language processing systems that deal with tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
+
+Because it is difficult to measure performance of a single (similarity) component in a pipeline, datasets that focus on similarity are popular among computational linguists.
+Note that the datasets not only measure potential utility of similarity models in a complex text processing system, but also serve as an empirical test of theories of Firth and Harris in their correspondence to psychological theories.
 
@@
 The datasets control for this property by providing similarity (relatedness) scores between word pairs.
+However, they overlook the following:
+\begin{compactitem}
+    % \item Similarity scores group words to categories that are formed around \emph{focal points}, the terms that represent clusters.
+    % To detect a focal point, one needs similarities between all words, or pairings between the focal point and the words of a category for each category.
+    \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
+    \item Global low similarity scores are ambiguous, they might mean dissimilarity between antonymy as between \textit{smart} and \textit{dumb} or absence of similarity in the case of \textit{trick} and \textit{size}.
+    % TODO: \cite{Roth1983346}
+    % TODO: this also leads to the extension of the similarity task: items are similar *because* of certain respect. Relevant respects have to be identified.
+\end{compactitem}
 %
-% TODO: the argument about focal points is still vague, because similarity scores still can express it.
-Maybe the argument should be given as a combination of centrality and ambiguity of low similarity.
-However, they overlook the following properties:
-\begin{compactitem}
-    \item Similarity scores group words to categories that are formed around \emph{focal points}, the terms that represent clusters.
-    % For this one needs to have all pairwise pairings or pairings between the focal point and the words of a category for each category.
-    \item Low similarity scores are ambiguous, they might mean dissimilarity between antonymy as between \textit{smart} and \textit{dumb} or absence of similarity in the case of \textit{trick} and \textit{size}.
-\end{compactitem}
+Consequently, we argue that a similarity task should be altered, so that a similarity model, instead of estimation of global similarity, provides similarity score together with a \emph{respect} the items are similar to.
 
@@
 
-\section{Similarity properties}
+\section{Similarity judgements}
 
-\paragraph{Centrality}
+% TODO: think of "respects". Goodman, Goldstone. Refer to Koen Lamberts & David Shanks "Knowledge, Concepts and Categories" page 51.
+% TODO: "Another possibility is that similarity and concepts are mutually constraining, and that neither presupposes the other." Lamberts&Shanks p. 66.
+
+Similarity is measured with respect to something \cite{}.
+It is very inaccurate (and probably impossible) to estimate the global similarity of ambiguous terms such as \textit{apple} and \textit{orange}.
+The two fruits are almost identical with respect to the shape, but have different taste.
+Both of them are used for juice production, thus have similar agricultural use.
+On the other hand, both of them are multinational companies, but they operate in different industries: technology and telecommunications.
+One could say that "\textit{apple}'s similarity to \textit{orange} is 0.8 because they are both textit{fruits}." and "\textit{apple}'s similarity to \textit{orange} is 0.6 because they are companies".
 
@@
 Where negative items indicate dissimilarity (hot, cold: -3), positive indicate similarity (hot, warm: 2) and 0 is reserved for incompatible pairs (trick, size:0)
-
+% * mutual constraint: similarity and concepts must be computed simultaneusly. H&C p. 83.
 % TODO: What if we ask not for similarity, but for confusion? Townsend (1971) "Theoretical analysis of an alphabetic confusion matrix".
@@
 
+% TODO: Sentence similarity.
+%
+%       How can we measure sentence similarity?
+%
+%         1) Head in context. E.g. "The boy runs." vs. "The girl walks."
+%         2) Full sentence simialrity. For example. "No" and "John took
+%            the cookies" are similar with the respect to the question 
+%            "Do we have cookies?"
+%         3) Transformational approach of either Kolmogorov complexity                  theory or "the number of changes between the referents".
+
 \bibliographystyle{acl2016}

### 2016-04-30 paper.tex

--- 2016-04-29 paper.tex
@@
 
-Similarity plays a crucial role in psychological theories of knowledge and behaviour, for example, it is used to classify objects \cite{Tversky1977}.
+Similarity plays a crucial role in psychological theories of knowledge and behaviour, for example, it is used to classify objects \cite{hahn1997concepts}.
 % TODO: maybe, mention \cite{Kriegeskorte2008}

### 2016-05-03 paper.tex
Tuesday

--- 2016-04-30 paper.tex
@@
 For example, it is used for classification and conceptualisation \cite{Tversky1977}.
-\textit{Fruit} is a \emph{category} because it is a useful generalisations in everyday life: fruits are sweet and constitute deserts, so when one is presented an unseen fruit, she can hypothesise that it is served towards the end of a dinner.
+\textit{Fruit} is a \emph{category} because it is a practical generalisation: Fruits are sweet and constitute deserts, so when one is presented an unseen fruit, she can hypothesise that it is served towards the end of a dinner.
 
@@
 
-Linguistic formalisation of similarity could base on two ideas: word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
+Linguistic formalisation of similarity bases on two ideas.
+Word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
 From computational perspective, this motivates and guides development of similarity components that are embedded into natural language processing systems that deal with tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
@@
 Because it is difficult to measure performance of a single (similarity) component in a pipeline, datasets that focus on similarity are popular among computational linguists.
-Note that the datasets not only measure potential utility of similarity models in a complex text processing system, but also serve as an empirical test of theories of Firth and Harris in their correspondence to psychological theories.
+Apart from a pragmatic attempt to alleviate the evaluation of similarity components, these datasets serve as an empirical test of theories of Firth and Harris, merging together our understanding of human mind, language and technology.
 
-Two, currently widely used, datasets are MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}.
-They are designed especially for meaning representation evaluation and surpass datasets stemmed in psychology\footnote{Refer to \newcite{1986-13502-00119860101} for a comprehensive list of datasets.} and information retrieval \cite{2002:PSC:503104.503110} in quantity and attention to the evaluated relation.
+Two datasets, namely MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}, are currently widely used.
+They are designed especially for meaning representation evaluation and surpass datasets stemmed in psychology \cite{1986-13502-00119860101} and information retrieval \cite{2002:PSC:503104.503110} in quantity and attention to the evaluated relation.
 
 Intuitively, similarity is the amount of common properties between entities, and can be implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
-The datasets control for this property by providing similarity (relatedness) scores between word pairs.
-However, they overlook the following:
+The datasets control for this property by providing similarity (relatedness) scores between word pairs, which provide \emph{global} similarity and overlook two aspects of similarity (relatedness):
 \begin{compactitem}
@@
     \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
-    \item Global low similarity scores are ambiguous, they might mean dissimilarity between antonymy as between \textit{smart} and \textit{dumb} or absence of similarity in the case of \textit{trick} and \textit{size}.
+    \item Similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
+For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
+
     % TODO: \cite{Roth1983346}
@@
 %
-Consequently, we argue that a similarity task should be altered, so that a similarity model, instead of estimation of global similarity, provides similarity score together with a \emph{respect} the items are similar to.
+Consequently, we argue that a similarity task should be altered, so that a similarity model, instead of estimation of global similarity, provides similarity score together with a \emph{respect} the items are similar to \cite{hahn1997concepts}.
 
@@
 
-\section{Similarity judgements}
+\section{Treating similarity with respect}
 
@@
 
-Similarity is measured with respect to something \cite{}.
+Similarity is measured with respect to something \cite{Roth1983346,Sadler1993}.
 It is very inaccurate (and probably impossible) to estimate the global similarity of ambiguous terms such as \textit{apple} and \textit{orange}.
@@
 On the other hand, both of them are multinational companies, but they operate in different industries: technology and telecommunications.
-One could say that "\textit{apple}'s similarity to \textit{orange} is 0.8 because they are both textit{fruits}." and "\textit{apple}'s similarity to \textit{orange} is 0.6 because they are companies".
+One could say that "\textit{apple}'s similarity to \textit{orange} is 8 out of 10 because they are both \textit{fruits}" and "as a company, \textit{apple}'s similarity to \textit{orange} is 6 out of 10." The context similarity is measured in is called \emph{respect} \cite{hahn1997concepts}.
 
-It has been shown empirically that similarity judgements geometrically arrange objects in such a way that the categories are formed around \emph{focal points}.
-This means that similarity judgements given by humans arrange fruit names around the word \textit{fruit} in such a way that it is their nearest neighbour, making \textit{fruit} the focal point of the category of \textit{fruits} \cite{1986-13502-00119860101}.
-This property is referred to as \emph{centrality}.
-%
-It is extremely important for a good semantic model to capture this property.
-Unfortunately, word pair similarity judgements do not capture this, as it is possible to arrange all fruit names close to \textit{fruit} but \textit{fruit} would not be the most common nearest neighbour.
+Once a representative number of similarity judgements is collected, a global similarity can be estimated by simply ignoring the respects.
+However, the respects themselves convey an important information.
+Hypothetically, a dominant respect between fruits is \textit{fruit}, cars is \textit{car} and so on.
+Thus, respects can be seen clusters, which similarity models need to identify.
+This solves the problem of ambiguity of low similarity scores, since there is not need to compare incompatible pairs.
+If there is no respect that ``connects'' two notions, this means that there is no similarity relation between them.
+% TODO: is it clear enough?
 
-To complicate the matter, existing lexical databases such as WordNet \cite{Miller:1995:WLD:219717.219748} can not be used to control for centrality.
-As \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies defined in those databases.
+Clustering is appealing because of another factor.
+If a category has a superordinate, similarity judgements arrange objects around it \cite{1986-13502-00119860101}.
+For example, similarity judgements given by humans arrange fruit names around the word \textit{fruit} in such a way that it is their nearest neighbour, making \textit{fruit} the \emph{focal point} of the category of \textit{fruits}.
+As an additional evaluation method, the model should be able to retrieve focal points.
+Note, however, that existing lexical databases such as WordNet \cite{Miller:1995:WLD:219717.219748} can not be used to recognise focal points.
+As \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies.
 For example, \textit{traffic} and \textit{water} might be considered to be similar because of functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
 
-\paragraph{Ambiguity}
+We propose to extend the similarity model evaluation to include classification, to overcome difficulties of uncontextualised similarity scores which might be due to the fact the similarity and concepts are computed simultaneously by humans \cite{hahn1997concepts}.
 
-Another issue is that the similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
-For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
+\section{Phrase and sentence similarity}
 
-While the presence of incompatible pairings can be motivated as an attempt to control for false positives, they sound unnatural to human.
-Also, a large amount of random pairings\footnotemark{} might influence a particular arrangement of the clusters of similar words, for example, the cluster of \textit{fruits} has to be between the cluster of \textit{furious adjectives} and \textit{mass nouns}, even if such arrangement was not intended by the dataset creators.
-\footnotetext{15\% of MEN entries have score less or equal to 10 out of 50 and by design 1000 word pairs are sampled from pairs that are assigned low cosine similarity scores by a text-based distributional model.}
-
-\section{Dataset design}
-
-% Some of the properties:
-%
-% * items belong to groups, such as fruits, animals, etc.
-% * all items in the group are assessed for similarity
-% * similarity is on the score of -3...+3.
-Where negative items indicate dissimilarity (hot, cold: -3), positive indicate similarity (hot, warm: 2) and 0 is reserved for incompatible pairs (trick, size:0)
-% * mutual constraint: similarity and concepts must be computed simultaneusly. H&C p. 83.
-% TODO: What if we ask not for similarity, but for confusion? Townsend (1971) "Theoretical analysis of an alphabetic confusion matrix".
-% TODO: Sorting seems to be a better fit! Rosenberg & Kim, (1975) "The Method of Sorting as a Data-Gathering Procedure in Multivariate Research".
-
-Evaluation methods need to focus on how well a model is able to recover human similarity intuitions expressed as groupings around their focal points (e.g~\textit{fruits}, \textit{companies} and \textit{mass nouns}).
-We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
-
-The similarity judgements should be collected in a different way which reflects human intuition about similarity.
-Even though vector space models of meaning measure similarity between words, it should not be (and can not) be contrasted directly with human judgements, instead the model needs to recover the concept clusters that humans define.
-
-% TODO: discuss antonymy and incompatible notions.
+In the previous section we illustrated similarity with examples between word pairs.
 
@@
 %         3) Transformational approach of either Kolmogorov complexity theory or "the number of changes between the referents".
+%         4) Truth conditional equivalence and possible words.
+
+\section{Conclustion}
+
+Similarity evaluation needs to focus on how well a model is able to recover human similarity intuitions expressed as groupings, possibly around their focal points.
+We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
 

### 2016-05-03 paper.tex
Sentences

@@
 
-In the previous section we illustrated similarity with examples between word pairs.
+In the previous section we focused on the properties of similarity, illustrating it with word-word examples.
+This section is dedicated to the nuances of similarity measurement between words, phrases and sentences.
+We assume that similarity requires the agreement of the grammatical roles of the words or phrases.
+
+A noun phrase can be similar to a noun, as in \textit{female lion} and \textit{lioness}, as between noun phrases with modifiers (\textit{yellow car} and \textit{taxi}), so the same similarity principles can be applied to phrases as to words.
+In this case, similarity is measured in \emph{context}.
+It is still comparison of head terms which meaning is modified by arguments they appear with \cite{Dinu:2010:MDS:1870658.1870771,mitchell2010composition}.
+With verbs this idea can be applied to compare transitive verbs with
+intransitive. For example, \textit{cycle} is similar to \textit{ride a bicycle}.
+
+Sentential similarity can be explained in several ways.
+First of all, it might be seen as the similarity of the heads in the contexts.
+That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}.
+This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found.
+
+Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but don not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid respects for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
 

### 2016-05-04 paper.tex
Wednesday
Minor changes

@@
 For example, it is used for classification and conceptualisation \cite{Tversky1977}.
-\textit{Fruit} is a \emph{category} because it is a practical generalisation: Fruits are sweet and constitute deserts, so when one is presented an unseen fruit, she can hypothesise that it is served towards the end of a dinner.
+\textit{Fruit} is a \emph{category} because it is a practical generalisation: Fruits are sweet and constitute deserts, so when one is presented with an unseen fruit, one can hypothesise that it is served towards the end of a dinner.
 
@@
 \textit{Verb}, \textit{subject} and \textit{singular} are categories that are used to describe English grammar.
-When one encounters an unknown word and is told that it is a verb, he will immediately have an idea how to use it assuming that it is used similarly to other English verbs.
+When one encounters an unknown word and is told that it is a verb, one will immediately have an idea how to use it assuming that it is used similarly to other English verbs.
 
-Linguistic formalisation of similarity bases on two ideas.
+Semantic formalisation of similarity bases on two ideas.
 Word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
@@
 Because it is difficult to measure performance of a single (similarity) component in a pipeline, datasets that focus on similarity are popular among computational linguists.
-Apart from a pragmatic attempt to alleviate the evaluation of similarity components, these datasets serve as an empirical test of theories of Firth and Harris, merging together our understanding of human mind, language and technology.
+Apart from a pragmatic attempt to alleviate the evaluation of similarity components, these datasets serve as an empirical test of the hypotheses of Firth and Harris, merging together our understanding of human mind, language and technology.
 
 Two datasets, namely MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}, are currently widely used.
-They are designed especially for meaning representation evaluation and surpass datasets stemmed in psychology \cite{1986-13502-00119860101} and information retrieval \cite{2002:PSC:503104.503110} in quantity and attention to the evaluated relation.
+They are designed especially for meaning representation evaluation and surpass datasets stemming from psychology \cite{1986-13502-00119860101} and information retrieval \cite{2002:PSC:503104.503110} in quantity by having more entries and, in case of SimLex-999, attention to the evaluated relation by distinguishing similarity from relatedness.
 
-Intuitively, similarity is the amount of common properties between entities, and can be implemented as a contrast model where ``the similarity between objects is expressed as a function of their common and distinctive features'' \cite{Tversky1977}.
-The datasets control for this property by providing similarity (relatedness) scores between word pairs, which provide \emph{global} similarity and overlook two aspects of similarity (relatedness):
+Similarity is the degree of resemblance between two objects or events \cite{WCS:WCS1282}.
+It can be implemented as a Contrast Model where the similarity between objects is expressed as a function of their common and distinctive features \cite{Tversky1977}.
+The datasets provide \emph{global} similarity (relatedness) scores between word pairs.
+Global similarity refers to an overall comparison \cite{hahn1997concepts} and is restricted with regard to the following observations:
 \begin{compactitem}
@@
     % To detect a focal point, one needs similarities between all words, or pairings between the focal point and the words of a category for each category.
-    \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
+    \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or kinds of fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
     \item Similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
@@
 On the other hand, both of them are multinational companies, but they operate in different industries: technology and telecommunications.
-One could say that "\textit{apple}'s similarity to \textit{orange} is 8 out of 10 because they are both \textit{fruits}" and "as a company, \textit{apple}'s similarity to \textit{orange} is 6 out of 10." The context similarity is measured in is called \emph{respect} \cite{hahn1997concepts}.
+One could say that ``\textit{apple}'s similarity to \textit{orange} is 8 out of 10 because they are both \textit{fruits}'' and ``as a company, \textit{apple}'s similarity to \textit{orange} is 6 out of 10.'' The context similarity is measured in is called \emph{respect} \cite{hahn1997concepts}.
 
@@
 Hypothetically, a dominant respect between fruits is \textit{fruit}, cars is \textit{car} and so on.
-Thus, respects can be seen clusters, which similarity models need to identify.
-This solves the problem of ambiguity of low similarity scores, since there is not need to compare incompatible pairs.
+Thus, respects can be seen as clusters, which similarity models need to identify.
+This solves the problem of ambiguity of low similarity scores, since there is no need to compare incompatible pairs.
 If there is no respect that ``connects'' two notions, this means that there is no similarity relation between them.
@@
 As an additional evaluation method, the model should be able to retrieve focal points.
-Note, however, that existing lexical databases such as WordNet \cite{Miller:1995:WLD:219717.219748} can not be used to recognise focal points.
+Note, however, that existing lexical databases such as WordNet \cite{Miller:1995:WLD:219717.219748} cannot be used to recognise focal points.
 As \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies.
-For example, \textit{traffic} and \textit{water} might be considered to be similar because of functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
+For example, \textit{traffic} and \textit{water} might be considered to be similar because of a functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
 
@@
 In this case, similarity is measured in \emph{context}.
-It is still comparison of head terms which meaning is modified by arguments they appear with \cite{Dinu:2010:MDS:1870658.1870771,mitchell2010composition}.
-With verbs this idea can be applied to compare transitive verbs with
-intransitive. For example, \textit{cycle} is similar to \textit{ride a bicycle}.
+It is still comparison of heads which meaning is modified by arguments they appear with \cite{Dinu:2010:MDS:1870658.1870771,mitchell2010composition}.
+With verbs this idea can be applied to compare transitive verbs with intransitive.
+For example, \textit{to cycle} is similar to \textit{ride a bicycle}.
 
@@
 
-Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but don not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid respects for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
+Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid respects for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
 

### 2016-05-04 paper.tex
an argument agains global similarity

@@
     % To detect a focal point, one needs similarities between all words, or pairings between the focal point and the words of a category for each category.
-    \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or kinds of fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
+    % TODO: the point should be against gathering global similarity, as most probably people will think of a respect when they give a similarity judgement.
+    \item ``[S]imilarity\ldots is a relationship between two objects \emph{under a given description}'' \cite{WCS:WCS1282}.
+Global similarity judgement omit the description and possibly give imprecise global similarity.
+Is \textit{apple}'s similarity to \textit{orange} as a company the same as a fruit?
     \item Similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).

### 2016-05-05 paper.tex
More on global similarity.
Intro

--- 2016-05-04 paper.tex
@@
 
-\title{On Similarity}
+\title{Treating Similarity with Respect: Recommendations on Evaluation of Models of Meaning}
 
@@
 
-Similarity plays a crucial role in psychological theories of knowledge and behaviour.
-For example, it is used for classification and conceptualisation \cite{Tversky1977}.
-\textit{Fruit} is a \emph{category} because it is a practical generalisation: Fruits are sweet and constitute deserts, so when one is presented with an unseen fruit, one can hypothesise that it is served towards the end of a dinner.
+Similarity is the degree of resemblance between two objects or events \cite{WCS:WCS1282} and plays a crucial role in psychological theories of knowledge and behaviour, where it is used to explain such phenomena as classification and conceptualisation.
+\textit{Fruit} is a \emph{category} because it is a practical generalisation.
+Fruits are sweet and constitute deserts, so when one is presented with an unseen fruit, one can hypothesise that it is served towards the end of a dinner.
 
@@
 They are designed especially for meaning representation evaluation and surpass datasets stemming from psychology \cite{1986-13502-00119860101} and information retrieval \cite{2002:PSC:503104.503110} in quantity by having more entries and, in case of SimLex-999, attention to the evaluated relation by distinguishing similarity from relatedness.
+The datasets provide \emph{global} similarity (relatedness) scores between word pairs.
 
-Similarity is the degree of resemblance between two objects or events \cite{WCS:WCS1282}.
-It can be implemented as a Contrast Model where the similarity between objects is expressed as a function of their common and distinctive features \cite{Tversky1977}.
-The datasets provide \emph{global} similarity (relatedness) scores between word pairs.
-Global similarity refers to an overall comparison \cite{hahn1997concepts} and is restricted with regard to the following observations:
-\begin{compactitem}
-    % \item Similarity scores group words to categories that are formed around \emph{focal points}, the terms that represent clusters.
-    % To detect a focal point, one needs similarities between all words, or pairings between the focal point and the words of a category for each category.
-    \item Similarity is measured with respect to something: \textit{apple} is similar to \textit{orange} because both are companies or kinds of fruits, but dissimilar because apples grow where oranges do not \cite{hahn1997concepts}.
-    \item Similarity scores presented by the datasets are ambiguous as it is not really clear what low similarity values mean: incompatible notions (\textit{trick} and \textit{size}) or contrast in meaning (\textit{smart} and \textit{dumb}).
-For example, SimLex-999 assigns low similarity scores to the pairs, 0.48 and 0.55 out of 10 respectively, but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
+In contrast to linguistic datasets, datasets that come from psychology contain entries that belong to a single category such as \textit{verbs of judging} \cite{FILLENBAUM197454}.
+This design decision might be motivated because similarity comparison seem to presuppose some amount or type of similarity.
+Moreover, because of the \emph{extension effect} \cite{medin1993respects}, the similarity of two entries in a context is less than the similarity between the same entries when the context is extended.
+``For example, \textit{black} and \textit{white} received a similarity rating of 2.2 when presented by themselves; this rating increased to 4.0 when \textit{black} was simultaneously compared with \textit{white} and \textit{red} (\textit{red} only increased 4.2 to 4.9).'' \cite{medin1993respects}.
+In the first case \textit{black} and \textit{white} are more dissimilar because they are located on the extremes of the greyscale, but in the presence of \textit{red} they become more similar because they are both monochromes.
 
-    % TODO: \cite{Roth1983346}
-    % TODO: this also leads to the extension of the similarity task: items are similar *because* of certain respect. Relevant respects have to be identified.
-\end{compactitem}
-%
-Consequently, we argue that a similarity task should be altered, so that a similarity model, instead of estimation of global similarity, provides similarity score together with a \emph{respect} the items are similar to \cite{hahn1997concepts}.
+Both MEN and SimLex provide pairs that do not share any similarity to control for false positives, and they do not control for the comparison scale.
+This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incompatible notions or contrast in meaning.
+SimLex-999 assigns low similarity scores to the incompatible pairs (0.48, \textit{trick} \textit{size}) and to antonymy (0.55, \textit{smart} and \textit{dumb}), but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
 
-% TODO: wrap up the introduction and make a transition to the next sections.
-
-\section{Treating similarity with respect}
+\section{Evaluation strategy}
 

### 2016-05-06 paper.tex
Minor changes

@@
 They are designed especially for meaning representation evaluation and surpass datasets stemming from psychology \cite{1986-13502-00119860101} and information retrieval \cite{2002:PSC:503104.503110} in quantity by having more entries and, in case of SimLex-999, attention to the evaluated relation by distinguishing similarity from relatedness.
-The datasets provide \emph{global} similarity (relatedness) scores between word pairs.
+The datasets provide similarity (relatedness) scores between word pairs.
 
-In contrast to linguistic datasets, datasets that come from psychology contain entries that belong to a single category such as \textit{verbs of judging} \cite{FILLENBAUM197454}.
+In contrast to linguistic datasets which contain randomly paired words from a broad selection, datasets that come from psychology contain entries that belong to a single category such as \textit{verbs of judging} \cite{FILLENBAUM197454}.
 This design decision might be motivated because similarity comparison seem to presuppose some amount or type of similarity.
 Moreover, because of the \emph{extension effect} \cite{medin1993respects}, the similarity of two entries in a context is less than the similarity between the same entries when the context is extended.
-``For example, \textit{black} and \textit{white} received a similarity rating of 2.2 when presented by themselves; this rating increased to 4.0 when \textit{black} was simultaneously compared with \textit{white} and \textit{red} (\textit{red} only increased 4.2 to 4.9).'' \cite{medin1993respects}.
+``For example, \textit{black} and \textit{white} received a similarity rating of 2.2 when presented by themselves; this rating increased to 4.0 when \textit{black} was simultaneously compared with \textit{white} and \textit{red} (\textit{red} only increased 4.2 to 4.9)'' \cite{medin1993respects}.
 In the first case \textit{black} and \textit{white} are more dissimilar because they are located on the extremes of the greyscale, but in the presence of \textit{red} they become more similar because they are both monochromes.

### 2016-05-06 paper.tex
The second page.

@@
 
-\title{Treating Similarity with Respect: Recommendations on Evaluation of Models of Meaning}
+\title{Treating Similarity with Respect: How to Evaluate Models of Meaning?}
 
@@
 \textit{Fruit} is a \emph{category} because it is a practical generalisation.
-Fruits are sweet and constitute deserts, so when one is presented with an unseen fruit, one can hypothesise that it is served towards the end of a dinner.
+Fruits are sweet and constitute deserts, so when one is presented with an unseen fruit, one can hypothesise that it is served toward the end of a dinner.
 
@@
 \textit{Verb}, \textit{subject} and \textit{singular} are categories that are used to describe English grammar.
-When one encounters an unknown word and is told that it is a verb, one will immediately have an idea how to use it assuming that it is used similarly to other English verbs.
+When one encounters an unknown word and is told that it is a verb, one will immediately have an idea about how to use it assuming that it is used similarly to other English verbs.
 
-Semantic formalisation of similarity bases on two ideas.
-Word occurrence patterns \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence \textup{quantifies} the difference in meaning \cite{harris1954distributional}.
-From computational perspective, this motivates and guides development of similarity components that are embedded into natural language processing systems that deal with tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
+The semantic formalisation of similarity is based on two ideas.
+The occurrence patterns of a word \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence between two words \emph{quantifies} the difference in their meaning \cite{harris1954distributional}.
+From a computational perspective, this motivates and guides development of similarity components that are embedded into natural language processing systems that deal with tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
 
 Because it is difficult to measure performance of a single (similarity) component in a pipeline, datasets that focus on similarity are popular among computational linguists.
-Apart from a pragmatic attempt to alleviate the evaluation of similarity components, these datasets serve as an empirical test of the hypotheses of Firth and Harris, merging together our understanding of human mind, language and technology.
+Apart from a pragmatic attempt to alleviate the problems of evaluating similarity components, these datasets serve as an empirical test of the hypotheses of Firth and Harris, bringing together our understanding of human mind, language and technology.
 
@@
 
-In contrast to linguistic datasets which contain randomly paired words from a broad selection, datasets that come from psychology contain entries that belong to a single category such as \textit{verbs of judging} \cite{FILLENBAUM197454}.
-This design decision might be motivated because similarity comparison seem to presuppose some amount or type of similarity.
+In contrast to linguistic datasets which contain randomly paired words from a broad selection, datasets that come from psychology contain entries that belong to a single category such as \textit{verbs of judging} \cite{FILLENBAUM197454} or \textit{animal terms} \cite{HENLEY1969176}.
+The reason for category oriented similarity studies is that ``stimuli can only be compared in so far as they have already been categorised as identical, alike, or equivalent at some higher level of abstraction'' \cite{turner1987rediscovering}.
 Moreover, because of the \emph{extension effect} \cite{medin1993respects}, the similarity of two entries in a context is less than the similarity between the same entries when the context is extended.
@@
 
-Both MEN and SimLex provide pairs that do not share any similarity to control for false positives, and they do not control for the comparison scale.
+Both MEN and SimLex-999 provide pairs that do not share any similarity to control for false positives, and they do not control for the comparison scale.
 This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incompatible notions or contrast in meaning.
@@
 
-\section{Evaluation strategy}
+This paper investigates how a similarity dataset with multiple categories should be built and considers what sentence similarity mean in this context.
 
-% TODO: think of "respects". Goodman, Goldstone. Refer to Koen Lamberts & David Shanks "Knowledge, Concepts and Categories" page 51.
-% TODO: "Another possibility is that similarity and concepts are mutually constraining, and that neither presupposes the other." Lamberts&Shanks p. 66.
+\section{Dataset construction and model evaluation strategies}
 
-Similarity is measured with respect to something \cite{Roth1983346,Sadler1993}.
-It is very inaccurate (and probably impossible) to estimate the global similarity of ambiguous terms such as \textit{apple} and \textit{orange}.
-The two fruits are almost identical with respect to the shape, but have different taste.
-Both of them are used for juice production, thus have similar agricultural use.
-On the other hand, both of them are multinational companies, but they operate in different industries: technology and telecommunications.
-One could say that ``\textit{apple}'s similarity to \textit{orange} is 8 out of 10 because they are both \textit{fruits}'' and ``as a company, \textit{apple}'s similarity to \textit{orange} is 6 out of 10.'' The context similarity is measured in is called \emph{respect} \cite{hahn1997concepts}.
+\paragraph{Human similarity judgements}
 
-Once a representative number of similarity judgements is collected, a global similarity can be estimated by simply ignoring the respects.
-However, the respects themselves convey an important information.
-Hypothetically, a dominant respect between fruits is \textit{fruit}, cars is \textit{car} and so on.
-Thus, respects can be seen as clusters, which similarity models need to identify.
-This solves the problem of ambiguity of low similarity scores, since there is no need to compare incompatible pairs.
-If there is no respect that ``connects'' two notions, this means that there is no similarity relation between them.
-% TODO: is it clear enough?
-
-Clustering is appealing because of another factor.
-If a category has a superordinate, similarity judgements arrange objects around it \cite{1986-13502-00119860101}.
-For example, similarity judgements given by humans arrange fruit names around the word \textit{fruit} in such a way that it is their nearest neighbour, making \textit{fruit} the \emph{focal point} of the category of \textit{fruits}.
-As an additional evaluation method, the model should be able to retrieve focal points.
-Note, however, that existing lexical databases such as WordNet \cite{Miller:1995:WLD:219717.219748} cannot be used to recognise focal points.
-As \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies.
+To build a similarity dataset that contains non-overlapping categories, one needs to avoid comparison of incompatible pairs.
+However, that itself requires a priori knowledge of item similarity or belongingness to a category, making the problem circular.
+To get out of the vicious circle, one might erroneously  refer to an already existing taxonomy such as WordNet \cite{Miller:1995:WLD:219717.219748}.
+But in case of similarity, as \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies.
 For example, \textit{traffic} and \textit{water} might be considered to be similar because of a functional similarity exploited in hydrodynamic models of traffic, but their lowest common ancestor in WordNet is \textit{entity}.
 
-We propose to extend the similarity model evaluation to include classification, to overcome difficulties of uncontextualised similarity scores which might be due to the fact the similarity and concepts are computed simultaneously by humans \cite{hahn1997concepts}.
+Since there is no way deciding upfront whether there is a similarity relation between two words, the data collection procedure needs test for both: relation existence and its strength.
+Numerical values, as has been shown in the introduction, do not fit this role, because of ambiguity.
+One way of avoiding the issue is to avoid asking humans for similarity judgements, but instead to ask them to list commonalities and differences between the objects.
+As one might expect, similarity scores correlate with the number of listed commonalities \cite{markman1991commonalities,Markman1996,medin1993respects}.
+For incompatible pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not another \cite{markman1991commonalities,medin1993respects}.
+
+\paragraph{The entries in the dataset}
+
+So far we have proposed a similarity judgement collection method that is robust to incompatible pairings.
+It also naturally gives rise to categories, because the absence of a relation between two entries means the absence of a common category.
+It still needs to be decided what words to include to the dataset.
+
+To get a list of words that constitute the dataset, one might think of categories such as \textit{sports}, \textit{fruits}, \textit{vegetatbles}, \textit{judging verbs}, \textit{countries}, \textit{colours} and so on.
+Note, that at this point its acceptable to think of categories, because later the arbitrary category assignments will be reevaluated by humans.
+Once the list of categories is ready, each of them is populated with category instances, e.g.~\textit{plum}, \textit{banana} ans \textit{lemon} are all \textit{fruits}.
+
+When the data is prepared, humans are asked to provide commonalities and differences between all pairs of every group.
+This serves two purposes.
+First, all expected similarities are judged, producing a dataset that can be seen as a merged version of category specific datasets.
+At this point, a good similarity model should provide meaning representation that are easily split to clusters: \textit{fruit} members and \textit{sport} members have to be separable.
+
+Intra-category comparisons should be also performed, but because it impractical to collect pairwise judgement between the number of words of magnitude of hundreds, a reasonable sample should be taken.
+The intra-category comparisons will lead to unexpected category pairings, such as \textit{food} that contains \textit{plants} and \textit{fruits}, so the sampling procedure might be directed by the discovery of compatible pairs: when a \textit{banana} and \textit{potato} are shred some commonalities, \textit{fruits} and \textit{vegetables} should be more likely to be assessed.
+
+\paragraph{Structure beyond proximity}
+
+Human judgements validate the initial category assignment of items and provide new ones.
+If a category contains a superordinate, similarity judgements arrange category members around it \cite{1986-13502-00119860101}.
+For example, similarity judgements given by humans arrange fruit names around the word \textit{fruit} in such a way that it is their nearest neighbour, making \textit{fruit} the \emph{focal point} of the category of \textit{fruits}.
+
+As an additional evaluation method, the model should be able to retrieve focal points.
+Therefore, a precaution should be taken before human judgement collection.
+If possible, categories should contain a subordinate.
+
+Similarity evaluation needs to focus on how well a model is able to recover human similarity intuitions expressed as groupings, possibly around their focal points.
+We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
+Models also should arrange subordinates to focal points giving labels to the clusters.
 
@@
 
-Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid respects for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
+Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid \textit{respects} for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
 
@@
 
-\section{Conclustion}
-
-Similarity evaluation needs to focus on how well a model is able to recover human similarity intuitions expressed as groupings, possibly around their focal points.
-We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
+\section{Conclusion}
 

### 2016-05-08 paper.tex
Submission

@@
 \usepackage{adjustbox}
+\usepackage{lingmacros}
 
@@
 % \aclfinalcopy % Uncomment this line for the final submission
-% \def\aclpaperid{34} %  Enter the acl Paper ID here
+\def\aclpaperid{19} %  Enter the acl Paper ID here
 
@@
 
-% \renewcommand{\baselinestretch}{0.95}
+\renewcommand{\baselinestretch}{0.95}
 
@@
 \begin{abstract}
-
+Similarity is a core notion that is used in psychology, theoretical and computational linguistics.
+The similarity datasets that come from the two fields differ in design: psychological datasets are focused around a certain topic such as fruit names; linguistic datasets contain words from various categories.
+The later makes humans assign low similarity scores to the words that have nothing in common and to the words that have contrast in meaning, making similarity scores ambiguous.
+In this work we discuss the similarity collection procedure for a multi-category dataset that avoids score ambiguity and suggest changes to the evaluation procedure to reflect the insights of psychological literature for word, phrase and sentence similarity.
+We suggest to ask humans to provide a list of commonalities and differences instead of numerical similarity scores and employ the structure of human judgements beyond pairwise similarity.
+We believe that the proposed approach will give rise to datasets that test meaning representation models more thoroughly with respect to the human treatment of similarity.
 \end{abstract}
@@
 The semantic formalisation of similarity is based on two ideas.
-The occurrence patterns of a word \emph{define} its meaning \cite{firth1957lingtheory}, while the difference in occurrence between two words \emph{quantifies} the difference in their meaning \cite{harris1954distributional}.
-From a computational perspective, this motivates and guides development of similarity components that are embedded into natural language processing systems that deal with tasks such as word sense disambiguation \cite{Schutze:1998:AWS:972719.972724}, information retrieval \cite{Salton:1975:VSM:361219.361220} and machine translation \cite{Dagan:1993:CWS:981574.981596}.
+The occurrence pattern of a word \emph{defines} its meaning \cite{firth1957lingtheory}, while the difference in occurrence between two words \emph{quantifies} the difference in their meaning \cite{harris1954distributional}.
+From a computational perspective, this motivates and guides development of similarity components that are embedded into natural language processing systems that deal with tasks such as
+word sense disambiguation \cite{Schutze:1998:AWS:972719.972724},
+information retrieval \cite{Salton:1975:VSM:361219.361220,Milajevs:2015:IMN:2808194.2809448},
+machine translation \cite{Dagan:1993:CWS:981574.981596},
+dependency parsing \cite{hermann-blunsom:2013:ACL2013,andreas-klein:2014:P14-2},
+and dialogue act tagging \cite{kalchbrenner-blunsom:2013:CVSC,milajevs-purver:2014:CVSC}.
+
+% TODO: Mention \cite{bengio2006}?
 
@@
 Two datasets, namely MEN \cite{Bruni:2012:DST:2390524.2390544} and SimLex-999 \cite{hill2014simlex}, are currently widely used.
-They are designed especially for meaning representation evaluation and surpass datasets stemming from psychology \cite{1986-13502-00119860101} and information retrieval \cite{2002:PSC:503104.503110} in quantity by having more entries and, in case of SimLex-999, attention to the evaluated relation by distinguishing similarity from relatedness.
+They are designed especially for meaning representation evaluation and surpass datasets stemming from psychology \cite{1986-13502-00119860101}, information retrieval \cite{2002:PSC:503104.503110} and computational linguistics \cite{Rubenstein:1965:CCS:365628.365657} in quantity by having more entries and, in case of SimLex-999, attention to the evaluated relation by distinguishing similarity from relatedness.
 The datasets provide similarity (relatedness) scores between word pairs.
@@
 This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incompatible notions or contrast in meaning.
-SimLex-999 assigns low similarity scores to the incompatible pairs (0.48, \textit{trick} \textit{size}) and to antonymy (0.55, \textit{smart} and \textit{dumb}), but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
+SimLex-999 assigns low similarity scores to the incompatible pairs (0.48, \textit{trick} and \textit{size}) and to antonymy (0.55, \textit{smart} and \textit{dumb}), but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
 
-This paper investigates how a similarity dataset with multiple categories should be built and considers what sentence similarity mean in this context.
+The present contribution investigates how a similarity dataset with multiple categories should be built and considers what sentence similarity means in this context.
 
-\section{Dataset construction and model evaluation strategies}
+\section{Dataset Construction}
 
@@
 
-To build a similarity dataset that contains non-overlapping categories, one needs to avoid comparison of incompatible pairs.
+To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incompatible pairs.
 However, that itself requires a priori knowledge of item similarity or belongingness to a category, making the problem circular.
-To get out of the vicious circle, one might erroneously  refer to an already existing taxonomy such as WordNet \cite{Miller:1995:WLD:219717.219748}.
+To get out of this vicious circle, one might erroneously  refer to an already existing taxonomy such as WordNet \cite{Miller:1995:WLD:219717.219748}.
 But in case of similarity, as \newcite{turney2012domain} points out, categories that emerge from similarity judgements are different from taxonomies.
@@
 
-Since there is no way deciding upfront whether there is a similarity relation between two words, the data collection procedure needs test for both: relation existence and its strength.
+Since there is no way of deciding upfront whether there is a similarity relation between two words, the data collection procedure needs to test for both: relation existence and its strength.
 Numerical values, as has been shown in the introduction, do not fit this role, because of ambiguity.
-One way of avoiding the issue is to avoid asking humans for similarity judgements, but instead to ask them to list commonalities and differences between the objects.
+One way of avoiding the issue is to avoid asking humans for numerical similarity judgements, but instead to ask them to list commonalities and differences between the objects.
 As one might expect, similarity scores correlate with the number of listed commonalities \cite{markman1991commonalities,Markman1996,medin1993respects}.
@@
 
-To get a list of words that constitute the dataset, one might think of categories such as \textit{sports}, \textit{fruits}, \textit{vegetatbles}, \textit{judging verbs}, \textit{countries}, \textit{colours} and so on.
-Note, that at this point its acceptable to think of categories, because later the arbitrary category assignments will be reevaluated by humans.
-Once the list of categories is ready, each of them is populated with category instances, e.g.~\textit{plum}, \textit{banana} ans \textit{lemon} are all \textit{fruits}.
+To get a list of words that constitute the dataset, one might think of categories such as \textit{sports}, \textit{fruits}, \textit{vegetables}, \textit{judging verbs}, \textit{countries}, \textit{colours} and so on.
+Note, that at this point its acceptable to think of categories, because later the arbitrary category assignments will be reevaluated.
+Once the list of categories is ready, each of them is populated with category instances, e.g.~\textit{plum}, \textit{banana} and \textit{lemon} are all \textit{fruits}.
 
 When the data is prepared, humans are asked to provide commonalities and differences between all pairs of every group.
-This serves two purposes.
 First, all expected similarities are judged, producing a dataset that can be seen as a merged version of category specific datasets.
@@
 
-Intra-category comparisons should be also performed, but because it impractical to collect pairwise judgement between the number of words of magnitude of hundreds, a reasonable sample should be taken.
-The intra-category comparisons will lead to unexpected category pairings, such as \textit{food} that contains \textit{plants} and \textit{fruits}, so the sampling procedure might be directed by the discovery of compatible pairs: when a \textit{banana} and \textit{potato} are shred some commonalities, \textit{fruits} and \textit{vegetables} should be more likely to be assessed.
+Intra-category comparisons should be also performed, but because it is impractical to collect all possible pairwise judgements between the number of words of magnitude of hundreds, a reasonable sample should be taken.
+The intra-category comparisons will lead to unexpected category pairings, such as \textit{food} that contains \textit{vegetables} and \textit{fruits}, so the sampling procedure might be directed by the discovery of compatible pairs: when a \textit{banana} and \textit{potato} are said to be similar, \textit{fruits} and \textit{vegetables} members should be more likely to be assessed.
 
-\paragraph{Structure beyond proximity}
+\paragraph{Evaluation beyond proximity}
 
@@
 
-\section{Phrase and sentence similarity}
+\section{Similarity in Context}
 
@@
 This section is dedicated to the nuances of similarity measurement between words, phrases and sentences.
-We assume that similarity requires the agreement of the grammatical roles of the words or phrases.
+Here, we assume that similarity requires the agreement of the grammatical roles.
 
-A noun phrase can be similar to a noun, as in \textit{female lion} and \textit{lioness}, as between noun phrases with modifiers (\textit{yellow car} and \textit{taxi}), so the same similarity principles can be applied to phrases as to words.
-In this case, similarity is measured in \emph{context}.
-It is still comparison of heads which meaning is modified by arguments they appear with \cite{Dinu:2010:MDS:1870658.1870771,mitchell2010composition}.
+A noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}.
+The same similarity principle can be applied to phrases as to words.
+In this case, similarity is measured in \emph{context}, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}.
 With verbs this idea can be applied to compare transitive verbs with intransitive.
@@
 
-Sentential similarity can be explained in several ways.
-First of all, it might be seen as the similarity of the heads in the contexts.
+% TODO: mention \cite{wieting2015paraphrase}
+
+Sentential similarity might be treated as the similarity of the heads in the contexts.
 That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}.
@@
 
-Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid \textit{respects} for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
+\section{Constructions and Language Sub-systems}
 
-% TODO: Sentence similarity.
-%
+The question of sentence similarity is more complex because sentences in many ways are different entities than words.
+Or are they?
+Linguistics has recently often pointed toward a continuum which exists between words and sentences \cite{jackendoff2012cambridge}.
+\newcite{jackendoff2005nature}, for example, point out that there is good evidence that ``human memory must store linguistic expressions of all sizes.'' These linguistic expressions of variable size are often called \emph{constructions}.
+Several computational approaches to constructions have been proposed \cite{Gaspers2011,chang2012computational}, but to the authors' best knowledge they do not yet feature prominently in natural language processing.
+
+Words, sentences and other constructions draw attention to states of affairs around us.
+This is both pragmatically and semantically speaking the case.
+As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{``X floosed Y the Z.''} and \textit{``X was floosed by Y.''} This is due to their similarity to sentences such as \textit{``John gave Mary the book.''} and \textit{``Mary was kissed by John.''} respectively.
+The point here is that some sentences are similar to others with respect to the functions they perform.
+For example, \textit{``X floosed Y the Z.''} is clearly a \emph{``transfer of possession"} or \emph{dative} \cite{bresnan2007predicting}.
+
+% As pointed out above, once one has categories one can speak of similarity.
+
+Thus, the amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence.
+\newcite{tomasello1998return} points out that sentence-level constructions show prototype effects similar to those discussed above for lexical systems (e.g.~colours).
+Consider the following sentences:
+\begin{compactitem}
+    \item \textit{``John gave Mary the book.''} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
+      \item \textit{``John promised Mary the book.''} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
+    %   \item \textit{``John denied Mary the book.''} is a example of an \emph{Agent Prevents Transfer} construction. These usually are build around words such as \textit{refuse, deny, etc.}
+    %   \item \textit{``John left Mary the book''} is a example of an \emph{Future Transfer} construction. These usually are build around words such as \textit{bequeath, leave, reserve, grant, etc.}
+    %   \item \textit{``John allowed Mary one book''} is a example of an \emph{Enabling Conditions for Transfer} construction. These usually are build around words such as \textit{permit, allow, etc.}
+    %   \item \textit{``John wrote Mary a book''} is a example of an \emph{Intended Transfer} construction.
+These usually are build around words such as \textit{bake, build, make, get, grab, win, earn, etc.}
+\end{compactitem}
+As soon as one has such a prototype network, one can actually decide sentence similarity as one can say with respect to what sentences and utterances are similar.
+
+However, these categories work on the semantic-grammatical level, and might be still handled by similarity in context as described in previous section.
+What about pragmatics?
+As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world.
+For example, he points to the fact that if a person asks another person to \textit{``pass the wine''} they are actually asking for the \textit{bottle}.
+The speaker just attracts attention to an object of perception in a given situation.
+There are several ways in which that sentence can both be \emph{grammaticalized} and \emph{lexicalized}.
+Dialogue act tags is another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
+%% Something about sentences with the same function in discourse here.
+
+If one conceptualises sentence similarity with respect to a discourse, then one might ask how different sentences fit in to such a discourse.
+\newcite{Griffiths2015a} tried to construct to versions of the same dialogue using a bottom-up method.
+They deconstructed a certain dialogue in a given domain---a receptionist scenario---into \textit{greetings}, \textit{directions} and \textit{farewells}.
+They used a small custom made corpus for this purpose and created the two dialogues by having people rate the individual utterances by friendliness.
+The resulting two dialogues were surprisingly uneven.
+The dialogue was supposed to give instructions to a certain location within a building.
+The ``friendly version'' was very elaborated and consisted of several sentences:
+\enumsentence{
+The questionnaire is located in room
+Q2-102. That is on the second floor. If
+you turn to your right and walk down
+the hallway. At the end of the floor you
+will find the stairs. Just walk up the
+stairs to the top floor and go through
+the fire door. The room is then straight
+ahead.}
+The sentence which served the same purpose in the neutral version was a fairly simple sentence:
+\enumsentence{
+The questionnaire is located in Q2-102.
+}
+Often the same functions in dialogue can be performed by as little as one word or several phrases or even a complete story.
+
+\newcite{Steels2010a} introduces the idea of language sub-systems and language strategies.
+A language subsystem is a means of expressing certain related or similar meanings.
+Examples of such sub-systems include:
+\begin{compactitem}
+    \item Lexical systems which express colours.
+    \item Morphological devices to encode tenses.
+    \item Using word order to express relations between agent and patient.
+\end{compactitem}  
+The later is an illustration of a language strategy.
+In English agent-patient relations are mainly encoded by syntax whereas German would use intonation and a combination or articles and case to convey the same information.
+Russian in contrast will use morphological devices for the same purpose.
+
+Also, sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{No} and \textit{I've seen John eating them} are similar sentences if they are answers to the question \textit{Do we have cookies?} Questions the sentences answer are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
+One approach to treat sentences on their own would be to encode the meaning of a sentence into a vector, in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.
+
+% We could also mention \cite{}.
+
 %       How can we measure sentence similarity?
 %
-%         1) Head in context. E.g. "The boy runs." vs. "The girl walks."
-%         2) Full sentence simialrity. For example. "No" and "John took
-%            the cookies" are similar with the respect to the question
-%            "Do we have cookies?"
 %         3) Transformational approach of either Kolmogorov complexity theory or "the number of changes between the referents".
@@
 
+In this contribution we discussed the notion of ``similarity'' from an interdisciplinary perspective.
+We contrasted properties of the similarity relation described in the field of psychology with the characteristics of similarity datasets used in computational linguistics.
+This lead to the recommendations on how to improve the later by removing low score ambiguity in a multi-category similarity dataset.
+
+In the future, a multi-category similarity dataset should be build that allows evaluation of vector space models of meaning by not only measuring proximity between the points, but also their arrangement to clusters.
+The same ideas can be used to build phrase and sentence level datasets.
+
+% At the very end, we need to say something about parsimony or "Occam's razor": Among competing hypotheses, the one with the fewest assumptions should be selected.
+%We are proposing going into a lot of detail in creating datasets and software. How do we leverage that with wide-coverage?
+
+On a broader perspective, this work highlights psychological phenomena that being incorporated into the models of meaning are expected to improve their performance.
+
+\balance
 \bibliographystyle{acl2016}
 \bibliography{references,dmilajevs_publications}
-\balance
 

### 2016-05-14 paper.tex
Sync
Getting ready for the final version

@@
 
-% \aclfinalcopy % Uncomment this line for the final submission
+\aclfinalcopy % Uncomment this line for the final submission
 \def\aclpaperid{19} %  Enter the acl Paper ID here
@@
 
-% \author{First Author \\
-%   Affiliation / Address line 1 \\
-%   Affiliation / Address line 2 \\
-%   Affiliation / Address line 3 \\
-%   {\tt email@domain} \\\And
-%   Second Author \\
-%   Affiliation / Address line 1 \\
-%   Affiliation / Address line 2 \\
-%   Affiliation / Address line 3 \\
-%   {\tt email@domain} \\}
+\author{Dmitrijs Milajevs \\
+  Queen Mary University of London \\
+  London, UK \\
+  {\tt d.milajevs@qmul.ac.uk} \\\And
+  Sascha Griffiths \\
+  Queen Mary University of London \\
+  London, UK \\
+  {\tt s.griffiths@qmul.ac.uk} \\
+}
 
@@
     \item Morphological devices to encode tenses.
-    \item Using word order to express relations between agent and patient.
+    \item Usage of word order to express relations between agent and patient.
 \end{compactitem}  

### 2016-06-13 paper.tex
comparable pairs and superordinates

@@
 Both MEN and SimLex-999 provide pairs that do not share any similarity to control for false positives, and they do not control for the comparison scale.
-This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incompatible notions or contrast in meaning.
-SimLex-999 assigns low similarity scores to the incompatible pairs (0.48, \textit{trick} and \textit{size}) and to antonymy (0.55, \textit{smart} and \textit{dumb}), but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
+This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incomparable notions or contrast in meaning.
+SimLex-999 assigns low similarity scores to the incomparable pairs (0.48, \textit{trick} and \textit{size}) and to antonymy (0.55, \textit{smart} and \textit{dumb}), but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
 
@@
 
-To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incompatible pairs.
+To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incomparable pairs.
 However, that itself requires a priori knowledge of item similarity or belongingness to a category, making the problem circular.
@@
 As one might expect, similarity scores correlate with the number of listed commonalities \cite{markman1991commonalities,Markman1996,medin1993respects}.
-For incompatible pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not another \cite{markman1991commonalities,medin1993respects}.
+For incomparable pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not another \cite{markman1991commonalities,medin1993respects}.
 
@@
 
-So far we have proposed a similarity judgement collection method that is robust to incompatible pairings.
+So far we have proposed a similarity judgement collection method that is robust to incomparable pairings.
 It also naturally gives rise to categories, because the absence of a relation between two entries means the absence of a common category.
@@
 Intra-category comparisons should be also performed, but because it is impractical to collect all possible pairwise judgements between the number of words of magnitude of hundreds, a reasonable sample should be taken.
-The intra-category comparisons will lead to unexpected category pairings, such as \textit{food} that contains \textit{vegetables} and \textit{fruits}, so the sampling procedure might be directed by the discovery of compatible pairs: when a \textit{banana} and \textit{potato} are said to be similar, \textit{fruits} and \textit{vegetables} members should be more likely to be assessed.
+The intra-category comparisons will lead to unexpected category pairings, such as \textit{food} that contains \textit{vegetables} and \textit{fruits}, so the sampling procedure might be directed by the discovery of comparable pairs: when a \textit{banana} and \textit{potato} are said to be similar, \textit{fruits} and \textit{vegetables} members should be more likely to be assessed.
 
@@
 Therefore, a precaution should be taken before human judgement collection.
-If possible, categories should contain a subordinate.
+If possible, categories should contain a superordinate.
 
@@
 We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
-Models also should arrange subordinates to focal points giving labels to the clusters.
+
+In contrast to the current evaluation based on correlation, models also need to be tested on geometric arrangement of subordinates around the focal points, as only the proximity based evaluation does not capture this \cite{1986-13502-00119860101}.
 

### 2016-06-13 paper.tex
Updates from ShareLaTeX

@@
 For incomparable pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not another \cite{markman1991commonalities,medin1993respects}.
+Verbally produced features (norms) for empirically derived conceptual representation of \newcite{McRae2005} is a good example of what data should be collected.
+But in contrast to \newcite{McRae2005}---where explicit comparison of concepts was avoided---participants should be asked to produce commonalities as part of similarity comparison.
 

### 2016-06-14 paper.tex
Gamification
Updates from ShareLaTeX

@@
 To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incomparable pairs.
-However, that itself requires a priori knowledge of item similarity or belongingness to a category, making the problem circular.
+However, that itself requires an a priori knowledge of item similarity or belongingness to a category, making the problem circular.
 To get out of this vicious circle, one might erroneously  refer to an already existing taxonomy such as WordNet \cite{Miller:1995:WLD:219717.219748}.
@@
 
+Given the dynamic nature of score collection, we suggest setting up \emph{a game with a purpose} (see \newcite{VenhuizenBasileEvangBos2013IWCS} for an example) where players are rewarded for contributing their commonality lists.
+
 \paragraph{Evaluation beyond proximity}
@@
 
-In contrast to the current evaluation based on correlation, models also need to be tested on geometric arrangement of subordinates around the focal points, as only the proximity based evaluation does not capture this \cite{1986-13502-00119860101}.
+In contrast to the current evaluation based on correlation, models also need to be tested on the geometric arrangement of subordinates around the focal points, as only the proximity based evaluation does not capture this \cite{1986-13502-00119860101}.
 
-\section{Similarity in Context}
+% \section{Similarity in Context}
 
-In the previous section we focused on the properties of similarity, illustrating it with word-word examples.
+% In the previous section we focused on the properties of similarity, illustrating it with word-word examples.
 This section is dedicated to the nuances of similarity measurement between words, phrases and sentences.
@@
 
-A noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}.
-The same similarity principle can be applied to phrases as to words.
-In this case, similarity is measured in \emph{context}, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}.
-With verbs this idea can be applied to compare transitive verbs with intransitive.
-For example, \textit{to cycle} is similar to \textit{ride a bicycle}.
-
-% TODO: mention \cite{wieting2015paraphrase}
-
-Sentential similarity might be treated as the similarity of the heads in the contexts.
-That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}.
-This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found.
-
-\section{Constructions and Language Sub-systems}
+% \section{Constructions and Language Sub-systems}
+\section{Sentence Similarity}
 
@@
 
+To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them. So what are they?
+
 Words, sentences and other constructions draw attention to states of affairs around us.
@@
 As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{``X floosed Y the Z.''} and \textit{``X was floosed by Y.''} This is due to their similarity to sentences such as \textit{``John gave Mary the book.''} and \textit{``Mary was kissed by John.''} respectively.
-The point here is that some sentences are similar to others with respect to the functions they perform.
+
+The point here is that some sentences are similar to others with respect to the \emph{functions} they perform.
 For example, \textit{``X floosed Y the Z.''} is clearly a \emph{``transfer of possession"} or \emph{dative} \cite{bresnan2007predicting}.
@@
 \end{compactitem}
+
 As soon as one has such a prototype network, one can actually decide sentence similarity as one can say with respect to what sentences and utterances are similar.
 
-However, these categories work on the semantic-grammatical level, and might be still handled by similarity in context as described in previous section.
-What about pragmatics?
+However, these categories work on the semantic-grammatical level, and might be handled by \emph{similarity in context}: a noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}.
+The same similarity principle can be applied to phrases as to words.
+In this case, similarity is measured in context, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}.
+With verbs this idea can be applied to compare transitive verbs with intransitive.
+For example, \textit{to cycle} is similar to \textit{to ride a bicycle}.
+
+Sentential similarity might be treated as the similarity of the heads in the contexts.
+That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}.
+This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found.
+
+% TODO: mention \cite{wieting2015paraphrase}
+
+But still, what about pragmatics?
 As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world.
@@
 The speaker just attracts attention to an object of perception in a given situation.
-There are several ways in which that sentence can both be \emph{grammaticalized} and \emph{lexicalized}.
-Dialogue act tags is another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
-%% Something about sentences with the same function in discourse here.
+
+There are several ways in which that sentence can both be \emph{grammaticalised} and \emph{lexicalised}.
+For example, \textit{No} and \textit{I've seen John eating them} are similar sentences because they lexicalise the same answer to the question \textit{Do we have cookies?} More generally, this gives rise to dialogue act tags: another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
+
+Thus, questions, which the sentences answer, are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
+This also motivates the approach of treating sentences on their own and encoding the meaning of a sentence into a vector in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.
 
 If one conceptualises sentence similarity with respect to a discourse, then one might ask how different sentences fit in to such a discourse.
-\newcite{Griffiths2015a} tried to construct to versions of the same dialogue using a bottom-up method.
+\newcite{Griffiths2015a} tried to construct two versions of the same dialogue using a bottom-up method.
 They deconstructed a certain dialogue in a given domain---a receptionist scenario---into \textit{greetings}, \textit{directions} and \textit{farewells}.
@@
 
-Also, sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but do not share grammatical structure at all, for example \textit{No} and \textit{I've seen John eating them} are similar sentences if they are answers to the question \textit{Do we have cookies?} Questions the sentences answer are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
-One approach to treat sentences on their own would be to encode the meaning of a sentence into a vector, in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.
-
-% We could also mention \cite{}.
-
-%       How can we measure sentence similarity?
-%
-%         3) Transformational approach of either Kolmogorov complexity theory or "the number of changes between the referents".
-%         4) Truth conditional equivalence and possible words.
-
 \section{Conclusion}

### 2016-06-14 paper.tex
New title and edits to the section on sentences.

@@
 
-\title{Treating Similarity with Respect: How to Evaluate Models of Meaning?}
+% \title{Treating Similarity with Respect: How to Evaluate Models of Meaning?}
+\title{A Proposal for Multi-category Lexical and Sentential Similarity Datasets Based on Commonality Lists}
 
@@
 
-% \section{Similarity in Context}
-
-% In the previous section we focused on the properties of similarity, illustrating it with word-word examples.
-This section is dedicated to the nuances of similarity measurement between words, phrases and sentences.
-Here, we assume that similarity requires the agreement of the grammatical roles.
-
-% \section{Constructions and Language Sub-systems}
 \section{Sentence Similarity}
@@
 
-To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them. So what are they?
+To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them.
+So what are they?
+First of all, words, sentences and other constructions draw attention to \emph{states of affairs} around us.
+% This is both pragmatically and semantically speaking the case.
 
-Words, sentences and other constructions draw attention to states of affairs around us.
-This is both pragmatically and semantically speaking the case.
-As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{``X floosed Y the Z.''} and \textit{``X was floosed by Y.''} This is due to their similarity to sentences such as \textit{``John gave Mary the book.''} and \textit{``Mary was kissed by John.''} respectively.
+\paragraph{Prototype effect}
 
-The point here is that some sentences are similar to others with respect to the \emph{functions} they perform.
-For example, \textit{``X floosed Y the Z.''} is clearly a \emph{``transfer of possession"} or \emph{dative} \cite{bresnan2007predicting}.
+Some sentences are similar to others with respect to the functions they perform.
+As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{X floosed Y the Z} and \textit{X was floosed by Y}.
+This is due to their similarity to sentences such as \textit{John gave Mary the book} and \textit{Mary was kissed by John} respectively.
+Thus, \textit{X floosed Y the Z} is clearly a \emph{transfer of possession} or \emph{dative} \cite{bresnan2007predicting}.
 
-% As pointed out above, once one has categories one can speak of similarity.
-
-Thus, the amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence.
-\newcite{tomasello1998return} points out that sentence-level constructions show prototype effects similar to those discussed above for lexical systems (e.g.~colours).
+The amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence.
+\newcite{tomasello1998return} points out that sentence-level constructions show \emph{prototype effects} similar to those discussed above for lexical systems (e.g.~colours).
 Consider the following sentences:
 \begin{compactitem}
-    \item \textit{``John gave Mary the book.''} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
-      \item \textit{``John promised Mary the book.''} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
+    \item \textit{John gave Mary the book.} is a example of an \emph{Agent Causes Transfer} construction. These usually are build around words such as \textit{give, pass, hand, toss, bring, etc.}
+      \item \textit{John promised Mary the book.} is a example of an \emph{Conditional transfer} construction. These usually are build around words such as \textit{promise, guarantee, owe, etc.}
     %   \item \textit{``John denied Mary the book.''} is a example of an \emph{Agent Prevents Transfer} construction. These usually are build around words such as \textit{refuse, deny, etc.}
@@
 
-However, these categories work on the semantic-grammatical level, and might be handled by \emph{similarity in context}: a noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}.
+\paragraph{Similarity in Context}
+
+However, prototype categories work on the semantic-grammatical level, and might be handled by \emph{similarity in context}: a noun phrase can be similar to a noun as in \textit{female lion} and \textit{lioness}, and to another noun phrases as in \textit{yellow car} and \textit{cheap taxi}.
 The same similarity principle can be applied to phrases as to words.
@@
 
+\paragraph{Attention Attraction}
+
 But still, what about pragmatics?
 As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world.
-For example, he points to the fact that if a person asks another person to \textit{``pass the wine''} they are actually asking for the \textit{bottle}.
+For example, he points to the fact that if a person asks another person to \textit{pass the wine} they are actually asking for the \textit{bottle}.
 The speaker just attracts attention to an object of perception in a given situation.
+
+\paragraph{Grammaticalisaton and Lexicalisaton}
 
@@
 This also motivates the approach of treating sentences on their own and encoding the meaning of a sentence into a vector in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.
+
+\paragraph{Discourse Fit}
 
@@
 
+\paragraph{Language Sub-systems and Strategies}
+
 \newcite{Steels2010a} introduces the idea of language sub-systems and language strategies.
@@
 
+Above we identified seven criteria by which sentence similarity can be compared.
+
 \section{Conclusion}
 
-In this contribution we discussed the notion of ``similarity'' from an interdisciplinary perspective.
+In this contribution we discussed the notion of similarity from an interdisciplinary perspective.
 We contrasted properties of the similarity relation described in the field of psychology with the characteristics of similarity datasets used in computational linguistics.

### 2016-06-14 paper.tex

@@
 % \title{Treating Similarity with Respect: How to Evaluate Models of Meaning?}
-\title{A Proposal for Multi-category Lexical and Sentential Similarity Datasets Based on Commonality Lists}
+\title{A Proposal for Linguistic Similarity Datasets Based on Commonality Lists}
 
@@
 \begin{abstract}
-Similarity is a core notion that is used in psychology, theoretical and computational linguistics.
-The similarity datasets that come from the two fields differ in design: psychological datasets are focused around a certain topic such as fruit names; linguistic datasets contain words from various categories.
+Similarity is a core notion that is used in psychology and two branches of linguistics: theoretical and computational.
+The similarity datasets that come from the two fields differ in design: psychological datasets are focused around a certain topic such as fruit names, while linguistic datasets contain words from various categories.
 The later makes humans assign low similarity scores to the words that have nothing in common and to the words that have contrast in meaning, making similarity scores ambiguous.
 In this work we discuss the similarity collection procedure for a multi-category dataset that avoids score ambiguity and suggest changes to the evaluation procedure to reflect the insights of psychological literature for word, phrase and sentence similarity.
-We suggest to ask humans to provide a list of commonalities and differences instead of numerical similarity scores and employ the structure of human judgements beyond pairwise similarity.
+We suggest to ask humans to provide a list of commonalities and differences instead of numerical similarity scores and employ the structure of human judgements beyond pairwise similarity for model evaluation.
 We believe that the proposed approach will give rise to datasets that test meaning representation models more thoroughly with respect to the human treatment of similarity.
@@
 \textit{Fruit} is a \emph{category} because it is a practical generalisation.
-Fruits are sweet and constitute deserts, so when one is presented with an unseen fruit, one can hypothesise that it is served toward the end of a dinner.
+Fruits are sweet and constitute deserts, so when one is presented with an unknown fruit, one can hypothesise that it is served toward the end of a dinner.
 
@@
 
-% TODO: Mention \cite{bengio2006}?
-
 Because it is difficult to measure performance of a single (similarity) component in a pipeline, datasets that focus on similarity are popular among computational linguists.
@@
 Both MEN and SimLex-999 provide pairs that do not share any similarity to control for false positives, and they do not control for the comparison scale.
-This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incomparable notions or contrast in meaning.
+This makes similarity judgements ambiguous as it is not clear what low similarity values mean: incomparable notions, contrast in meaning or even the difference in comparison context.
 SimLex-999 assigns low similarity scores to the incomparable pairs (0.48, \textit{trick} and \textit{size}) and to antonymy (0.55, \textit{smart} and \textit{dumb}), but \textit{smart} and \textit{dumb} have relatively much more in common than \textit{trick} and \textit{size}!
@@
 
-To build a similarity dataset which contains non-overlapping categories, one needs to avoid comparison of incomparable pairs.
+To build a similarity dataset that contains non-overlapping categories, one needs to avoid comparison of incomparable pairs.
 However, that itself requires an a priori knowledge of item similarity or belongingness to a category, making the problem circular.
@@
 For incomparable pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not another \cite{markman1991commonalities,medin1993respects}.
-Verbally produced features (norms) for empirically derived conceptual representation of \newcite{McRae2005} is a good example of what data should be collected.
+Verbally produced features (norms) for empirically derived conceptual representation of \newcite{McRae2005} is a good example of what and how the data should be collected.
 But in contrast to \newcite{McRae2005}---where explicit comparison of concepts was avoided---participants should be asked to produce commonalities as part of similarity comparison.
@@
 
-So far we have proposed a similarity judgement collection method that is robust to incomparable pairings.
+So far, we have proposed a similarity judgement collection method that is robust to incomparable pairings.
 It also naturally gives rise to categories, because the absence of a relation between two entries means the absence of a common category.
@@
 
-\paragraph{Prototype effect}
+\paragraph{Prototype effects}
 
@@
 
-\paragraph{Similarity in Context}
+\paragraph{Similarity in context}
 
@@
 
-\paragraph{Attention Attraction}
+\paragraph{Attention attraction}
 
@@
 
-\paragraph{Grammaticalisaton and Lexicalisaton}
+\paragraph{Grammaticalisaton and lexicalisaton}
 
@@
 
-\paragraph{Discourse Fit}
+\paragraph{Discourse fit}
 
@@
 
-\paragraph{Language Sub-systems and Strategies}
+\paragraph{Language Sub-systems and strategies}
 

### 2016-06-14 paper.tex
Minor changes

@@
 That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}.
-This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found.
+This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found \cite{corbett1993heads}.
 
@@
 
-There are several ways in which that sentence can both be \emph{grammaticalised} and \emph{lexicalised}.
+There are several ways in which a sentence can both be \emph{grammaticalised} and \emph{lexicalised}.
 For example, \textit{No} and \textit{I've seen John eating them} are similar sentences because they lexicalise the same answer to the question \textit{Do we have cookies?} More generally, this gives rise to dialogue act tags: another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
@@
 
-\paragraph{Language Sub-systems and strategies}
+\paragraph{Language sub-systems and strategies}
 
 \newcite{Steels2010a} introduces the idea of language sub-systems and language strategies.
-A language subsystem is a means of expressing certain related or similar meanings.
+A language subsystem are the means of expressing certain related or similar meanings.
 Examples of such sub-systems include:
@@
 In English agent-patient relations are mainly encoded by syntax whereas German would use intonation and a combination or articles and case to convey the same information.
-Russian in contrast will use morphological devices for the same purpose.
+Russian, in contrast, will use morphological devices for the same purpose.
 
@@
 In the future, a multi-category similarity dataset should be build that allows evaluation of vector space models of meaning by not only measuring proximity between the points, but also their arrangement to clusters.
-The same ideas can be used to build phrase and sentence level datasets.
-
-% At the very end, we need to say something about parsimony or "Occam's razor": Among competing hypotheses, the one with the fewest assumptions should be selected.
-%We are proposing going into a lot of detail in creating datasets and software. How do we leverage that with wide-coverage?
+The same ideas can be used to build phrase- and sentence-level datasets.
 

### 2016-06-17 paper.tex

@@
 Since there is no way of deciding upfront whether there is a similarity relation between two words, the data collection procedure needs to test for both: relation existence and its strength.
-Numerical values, as has been shown in the introduction, do not fit this role, because of ambiguity.
-One way of avoiding the issue is to avoid asking humans for numerical similarity judgements, but instead to ask them to list commonalities and differences between the objects.
+Numerical values, as has been shown in the introduction, do not fit this role due to ambiguity.
+One way to avoid the issue is to avoid asking humans for numerical similarity judgements, but instead to ask them to list commonalities and differences between the objects.
 As one might expect, similarity scores correlate with the number of listed commonalities \cite{markman1991commonalities,Markman1996,medin1993respects}.
-For incomparable pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not another \cite{markman1991commonalities,medin1993respects}.
+For incomparable pairs, the commonality list should be empty, but the differences will enumerate properties that belong to one entity, but not to another \cite{markman1991commonalities,medin1993respects}.
+
 Verbally produced features (norms) for empirically derived conceptual representation of \newcite{McRae2005} is a good example of what and how the data should be collected.
@@
 It also naturally gives rise to categories, because the absence of a relation between two entries means the absence of a common category.
-It still needs to be decided what words to include to the dataset.
+It still needs to be decided which words to include in the dataset.
 
@@
 
-Given the dynamic nature of score collection, we suggest setting up \emph{a game with a purpose} (see \newcite{VenhuizenBasileEvangBos2013IWCS} for an example) where players are rewarded for contributing their commonality lists.
+Given the dynamic nature of score collection, we suggest setting up \emph{a game with a purpose} (see \newcite{VenhuizenBasileEvangBos2013IWCS} an example) where players are rewarded for contributing their commonality lists.
+Another option would be to crowdsource the human judgements \cite{doi:10.1080/17470218.2015.1051065}.
 
@@
 Similarity evaluation needs to focus on how well a model is able to recover human similarity intuitions expressed as groupings, possibly around their focal points.
-We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to a same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than a \textit{company}.
+We propose to treat it as a soft multi-class clustering problem \cite{White:2015:WSE:2838931.2838932}, where two entities belong to the same class if there is a similarity judgement for them (e.g.~\textit{apple} and \textit{banana} are similar because they are \textit{fruits}) and the strength is proportional to the number of such judgements, so we could express that \textit{apple} is more a \textit{fruit} than it is a \textit{company}.
 
@@
 
-To be able to measure similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them.
+To be able to measure the similarity of phrases and sentences in the proposed framework, we need to be able to identify what could serve as commonalities between them.
 So what are they?
 First of all, words, sentences and other constructions draw attention to \emph{states of affairs} around us.
+Also, sentences are similar to others with respect to the functions they perform \cite[p.
+288]{Winograd1983}.
 % This is both pragmatically and semantically speaking the case.
@@
 
-Some sentences are similar to others with respect to the functions they perform.
 As \newcite{tomasello2009cultural} points out, speakers of English can make sense of phrases like \textit{X floosed Y the Z} and \textit{X was floosed by Y}.
@@
 
-The amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence.
+The amount in which sentences are similar, at least to a certain extent, corresponds to the function of a given sentence (the ideational function \cite[p.
+288]{Winograd1983} especially).
 \newcite{tomasello1998return} points out that sentence-level constructions show \emph{prototype effects} similar to those discussed above for lexical systems (e.g.~colours).
@@
 
-As soon as one has such a prototype network, one can actually decide sentence similarity as one can say with respect to what sentences and utterances are similar.
+As soon as one has such a prototype network, one can actually decide sentence similarity as one can say with respect to what prototypes sentences and utterances are similar.
+In this case, a common sentence prototype serves the same role as commonality between words.
 
@@
 The same similarity principle can be applied to phrases as to words.
-In this case, similarity is measured in context, but is still comparison of the phrases' head words which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}.
+In this case, similarity is measured in context, but it is still a comparison of the phrases' head words of which meaning is modified by arguments they appear with \cite{Kintsch2001173,mitchell-lapata:2008:ACLMain,mitchell2010composition,Dinu:2010:MDS:1870658.1870771,Baroni2010nouns,thater-furstenau-pinkal:2011:IJCNLP-2011,Seaghdha:2011:PMS:2145432.2145545}.
 With verbs this idea can be applied to compare transitive verbs with intransitive.
@@
 That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}.
-This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found \cite{corbett1993heads}.
+This approach abstracts away from grammatical differences between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found \cite{corbett1993heads}.
 
@@
 But still, what about pragmatics?
-As \newcite{Steels2008} points out, sentences and words draw attention and do not always directly point or refer to entities and actions in the world.
+As \newcite{Steels2008} points out, sentences and words direct attention and do not always directly point or refer to entities and actions in the world.
 For example, he points to the fact that if a person asks another person to \textit{pass the wine} they are actually asking for the \textit{bottle}.
@@
 There are several ways in which a sentence can both be \emph{grammaticalised} and \emph{lexicalised}.
-For example, \textit{No} and \textit{I've seen John eating them} are similar sentences because they lexicalise the same answer to the question \textit{Do we have cookies?} More generally, this gives rise to dialogue act tags: another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
+For example, \textit{No} and \textit{I've seen John eating them} are similar sentences because they lexicalise the same answer to the question \textit{Do we have cookies?} More generally, this gives rise to dialogue act tags: for another way of utterance categorisation, refer to the work of \newcite{kalchbrenner-blunsom:2013:CVSC} and \newcite{milajevs-purver:2014:CVSC}.
 
-Thus, questions, which the sentences answer, are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
+Thus, questions which the sentences answer, are valid respects for similarity explanation, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
 This also motivates the approach of treating sentences on their own and encoding the meaning of a sentence into a vector in such a way that similar sentences are clustered together \cite{DBLP:journals/corr/abs-1003-4394,baroni2014frege,Socher:2012:SCT:2390948.2391084,wieting2015paraphrase,DBLP:journals/corr/HillCK16}.
@@
 ahead.}
-The sentence which served the same purpose in the neutral version was a fairly simple sentence:
+The sentence which served the same purpose in the ``neutral version'' was a fairly simple sentence:
 \enumsentence{
@@
 }
-Often the same functions in dialogue can be performed by as little as one word or several phrases or even a complete story.
+Often the same function of a given sentence in a dialogue can be performed by as little as one word or several phrases or a different sentence or even a complete story.
 
@@
 Russian, in contrast, will use morphological devices for the same purpose.
+Hence, for some purposes the entities which are similar may not be of clearly delineated categories such as ``word'' or ``sentence'' but may be of chunks of language which belong to the same sub-system.
 
 Above we identified seven criteria by which sentence similarity can be compared.
+The instructions for the sentence similarity judgement tasks may incorporate the criteria as hints for human participants during data collection.
 
@@
 
-In the future, a multi-category similarity dataset should be build that allows evaluation of vector space models of meaning by not only measuring proximity between the points, but also their arrangement to clusters.
+In the future, a multi-category similarity dataset should be build that allow evaluation of vector space models of meaning by not only measuring proximity between the points, but also their arrangement with respect to clusters.
 The same ideas can be used to build phrase- and sentence-level datasets.
+However, we leave the exact sentence similarity criteria selection for future work in this area.
 
 On a broader perspective, this work highlights psychological phenomena that being incorporated into the models of meaning are expected to improve their performance.
+
+\section*{Acknowledgements}
+
+We thank the anonymous reviewers for their comments. Support from EPSRC grant EP/J002607/1 is gratefully acknowledged by Dmitrijs Milajevs.
+Sascha Griffiths is supported by ConCreTe: the project ConCreTe acknowledges the financial support of the Future and Emerging Technologies (FET) programme within the Seventh Framework Programme for Research of the European Commission, under FET grant number 611733.
 
